---
title: "Load and process data from Zhan et al. 2006"
author: "TYM"
date: "`r format(Sys.time(), '%d-%m-%Y')`, Aalborg, Denmark"
output: html_document
---

Load packages.
```{r}
library(affy)
library(hgu133plus2cdf)
library(GEOquery)
library(magrittr)
library(tidyverse)
library(arrangements)
library(mclust)

# Reordering a confusion matrix to optimize accuracy
opt <- function(x){
  acc <- function(x)
  {
    sum(diag(x))/sum(x)
  }
  pe <- permutations(7,7)
  curr.x <- acc(x)
  curr.i <- 1:7
  for(i in 1:nrow(pe)){
    if(acc(x[pe[i,],])>curr.x){ 
      curr.i = pe[i,]
      curr.x <- acc(x[pe[i,],])
    }
  }
  curr.i
}
```
Download the raw .CEL files from Zhan et al. 2006.
```{r}
getGEOSuppFiles("GSE24080",baseDir = "ExternalData/MAQCII_GSE24080",makeDirectory = FALSE)
```

Decompress the data and delete compressed data.
```{bash}
# .CEL files. 
tar -xvf ExternalData/MAQCII_GSE24080/GSE24080_RAW.tar -C ExternalData/MAQCII_GSE24080/ 
rm ExternalData/MAQCII_GSE24080/GSE24080_RAW.tar
```

Load the preprocessed GSE datasets to cross reference IDs.
```{r}
# Load the two datasets.
if(!file.exists("ExternalData/UAMS_GSE4581/gse4581.RData")){
  gse4581 <- getGEO("GSE4581", GSEMatrix = TRUE,destdir = "ExternalData/UAMS_GSE4581")[[1]]
  save(gse4581, file = "ExternalData/UAMS_GSE4581/gse4581.RData")
}
load("ExternalData/UAMS_GSE4581/gse4581.RData")

if(!file.exists("ExternalData/MAQCII_GSE24080/gse24080.RData")){
  gse24080 <- getGEO("GSE24080", GSEMatrix = TRUE,destdir = "ExternalData/MAQCII_GSE24080")[[1]]
  save(gse24080,file = "ExternalData/MAQCII_GSE24080/gse24080.RData")
}
load("ExternalData/MAQCII_GSE24080/gse24080.RData")

# Cross reference the IDs.
meta_4581  <- pData(gse4581) %>%
  mutate(
    PatientID = gsub("U133Plus-","",title) %>% 
                gsub(" \\(.*","",.) %>% 
                gsub("-.*","",.)) 

meta_24080 <- pData(gse24080) %>%
  mutate(
    PatientID = gsub("-.*","",title) %>%
    sapply(.,function(i){
      Str <- strsplit(i,split = "")[[1]]
      
      if (Str[2] == "0"){
        out <- paste(Str[-2],collapse = "")
      } else {
        out <- paste(Str,collapse = "")
      }
      return(out)
    }) %>% as.character())

# Merge the two IDs.
df_merg    <- left_join(select(meta_4581,PatientID,geo_accession),
                        select(meta_24080,PatientID,geo_accession),
                        by = "PatientID",suffix = c("_4581","_24080"))

# Find the filenames matching. Note that only 408 of 414 had a match.
all_files  <- list.files("ExternalData/MAQCII_GSE24080",pattern = "CEL.gz",full.names = T)
files_zhan <- df_merg$geo_accession_24080 %>%
  sapply(.,function(i){
    tmp <- grep(pattern = i,x = all_files,value = T)
    ifelse(all(is.na(tmp)),NA,tmp)
  }) %>%
  {.[!is.na(.)]}

# Test that it is the same data!
wh <- names(files_zhan) %>% sapply(.,function(i){which(i == df_merg$geo_accession_24080)})

x1 <- log2(exprs(gse4581))[,df_merg$geo_accession_4581[wh]]
x2 <- exprs(gse24080)[,df_merg$geo_accession_24080[wh]]

cors <- mapply(function(a,b){
  cor(x1[,a],x2[,b],use = "complete.obs")
},a = colnames(x1),b = colnames(x2)) %>% `names<-`(colnames(x2))
  
boxplot(cors) # All but one sample have r > 0.9999

# Remove the file with bad correlation.
bad        <- names(which.min(cors))
files_zhan <- files_zhan[!grepl(bad,files_zhan)]
```

Load the raw data from zhan et al. 2006.
```{r}
if(!file.exists("GeneratedData/Zhan_mas5.rds")){

  # Read the CEL files into R.
  zhan_data <- read.affybatch(files_zhan,compress = T)

  # MAS5 normalization.
  zhan_mas5 <- mas5(zhan_data)
  
  # Present/absent calls.
  zhan_PA   <- mas5calls(zhan_data)
  
  # Save relevant data.
  saveRDS(zhan_mas5,"GeneratedData/Zhan_mas5.rds")
  saveRDS(zhan_PA,"GeneratedData/Zhan_PA.rds")
}
zhan_mas5 <- readRDS("GeneratedData/Zhan_mas5.rds")
zhan_PA   <- readRDS("GeneratedData/Zhan_PA.rds")
````

Curate the metadata from Zhan et al. 2006.
```{r}
pheno.mm <- pData(gse4581)

## Get treatment regime

pheno.mm$regime <- "TT2"
pheno.mm$regime[grep("TT3", pheno.mm$title)] <- "TT3"

## Get the eventtime, event, and class phenotype and removing annoying extra chars

pheno.mm$eventtime <- as.character(pheno.mm$characteristics_ch1.2)
pheno.mm$eventtime <- sub("\\[SURTIM=", "", pheno.mm$eventtime)
pheno.mm$eventtime <- as.numeric(lapply(strsplit(pheno.mm$eventtime, " "),
                                        function(x){x[1]}))

pheno.mm$event <- as.character(pheno.mm$characteristics_ch1)
pheno.mm$event <- sub("\\[SURIND=", "", pheno.mm$event)
pheno.mm$event <- as.numeric(lapply(strsplit(pheno.mm$event, " "),
                                        function(x){x[1]}))

pheno.mm$class <-  pData(gse4581)$characteristics_ch1.8
pheno.mm$class <- sub("\\[Subgrp7=", "", pheno.mm$class)
pheno.mm$class <- factor(sub("\\]", "", pheno.mm$class))
pheno.mm$class.no <- as.numeric(pheno.mm$class) 

pheno.mm$PatientID <- gsub("U133Plus-","",pheno.mm$title) %>% 
                      gsub(" \\(.*","",.) %>% 
                      gsub("-.*","",.)
pheno.mm <- select(pheno.mm,PatientID,regime,eventtime,event,class,class.no)
```

Merge the phenodata with expression data, such that IDs match.
```{r}
CELtoPatientID <- pData(zhan_mas5) %>% rownames_to_column("CEL") %>%
  mutate(
    sample = gsub("-.*","",CEL) %>%
    sapply(.,function(i){strsplit(i,split = "_")[[1]][2]}) %>%
    sapply(.,function(i){
      Str <- strsplit(i,split = "")[[1]]
      
      if (Str[2] == "0"){
        out <- paste(Str[-2],collapse = "")
      } else {
        out <- paste(Str,collapse = "")
      }
      return(out)
    }) %>% as.character())

zhan_pheno <- left_join(x,pheno.mm,by = c("sample" = "PatientID")) %>% column_to_rownames("CEL")

pData(zhan_mas5) <- zhan_pheno
fData(zhan_mas5) <- fData(gse4581)
```

Now perform the filtering according to Zhan et al. 2006.
```{r}
# 1. Keep probesets called present in >3% of samples.---------------------------
keeps <- exprs(zhan_PA) %>% 
  apply(.,1,`==`,"P") %>% 
  apply(.,2,function(x){sum(x) > floor(length(x)*0.03)})

zhan_proc <- zhan_mas5[keeps,]

# 2. log2 transform.------------------------------------------------------------
exprs(zhan_proc) <- log2(exprs(zhan_proc))

# 3. Retain only affymetrix probesets with a standard deviation > 1.34 in the training dataset.
exprs.train <- exprs(zhan_proc[,zhan_proc$regime == "TT2"])
keeps       <- apply(exprs.train,1,sd) > 1.34

zhan_proc   <- zhan_proc[keeps,]

# 4. For multiple affymetrix probesets matching the same gene, retain the one with highest variance.

# Get the set of all gene symbols.
geneIDs <- fData(zhan_proc)$`Gene Symbol` %>%
  lapply(.,function(x){strsplit(x,split = " /// ")[[1]]}) %>%
  unlist() %>%
  unique()
geneIDs <- unique(fData(zhan_proc)$`Gene Symbol`)

# Go through each gene iteratively.
tmp   <- fData(zhan_proc)$`Gene Symbol`
keeps <- sapply(geneIDs,function(gene){
  wh    <- which(tmp == gene)
  x_sub <- exprs(zhan_proc)[wh,,drop = F]
  names(which.max(apply(x_sub,1,sd)))
})

zhan_proc <- zhan_proc[keeps,]
```

Save the filtered data.
```{r}
saveRDS(zhan_proc,"GeneratedData/Zhan_filtered.rds")
```

# Check that the data is similar to GSE4581 

```{r}
wh_probes <- intersect(rownames(zhan_proc),rownames(gse4581))

df_merg <- pData(zhan_proc) %>% rownames_to_column("CEL") %>%
  left_join(.,meta_4581,by = c("sample" = "PatientID")) %>%
  select(CEL,geo_accession,sample,title)

cors <- sapply(1:nrow(df_merg),function(i){
  cor(exprs(zhan_proc)[wh_probes,df_merg$CEL[i]],log2(exprs(gse4581)[wh_probes,df_merg$geo_accession[i]]),use = "complete.obs")
})
boxplot(cors)
```

# Hierarchical clustering 

We try to cluster the data and cut at 7 groups to see if we get the same results as Zhan 2006.
```{r}
d <- dist(t(exprs(zhan_proc[,zhan_proc$regime == "TT2"])))

clust <- hclust(d)

labs <- cutree(clust,k = 7) %>% data.frame(clust = .)

tmp <- merge(pData(zhan_proc),labs,by = 0)

conf.GMM <- table(tmp$class.no,tmp$clust)

# Reorder to get "diagonal" matrix
conf.hclst <- conf.hclst[opt(conf.hclst), ]
conf.hclst
sum(diag(conf.hclst))/sum(conf.hclst)
```

# GMM clustering

We try to cluster the data and cut at 7 groups to see if we get the same results as Zhan 2006.
```{r}
fit <- Mclust(t(exprs(zhan_proc[,zhan_proc$regime == "TT2"])),G = 7)

labs <- fit$classification %>% data.frame(clust = .)

tmp <- merge(pData(zhan_proc),labs,by = 0)

conf.GMM <- table(tmp$class.no,tmp$clust)

# Reorder to get "diagonal" matrix
conf.GMM <- conf.GMM[opt(conf.GMM), ]
conf.GMM
sum(diag(conf.GMM))/sum(conf.GMM)
```

